---
title: "BEAST with spatial data"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Load needed libraries. 

Install if necessary.

```{r}

#  convoluted leftovers from initial version
#  should just load tbreak directly
pack_list = c("tbreak", "sf")
installed = installed.packages()
install_list = pack_list[!(pack_list %in% installed[,"Package"])]
if(length(install_list) || installed[,"Version"][["tbreak"]] < "0.7") {
  if (!("remotes" %in% installed[,"Package"])) {
    install.packages("remotes")
  }
  library ("remotes")
  install_github("biogeospatial/tbreak")
}

for (p in pack_list) {
  library(p, character.only = TRUE)
}

#  Other libs needed below
pack_list = c("svDialogs", "terra", "stringi")
install_list = pack_list[!(pack_list %in% installed.packages()[,"Package"])]
if(length(install_list)) {
  install.packages(install_list)
}
for (p in pack_list) {
  library(p, character.only = TRUE)
}


```

#  Load the raster 

Now load the file we want and display some of its properties, and the
first band.

Make sure you load a geotiff or netcdf file with *ONLY ONE INDEX*, e.g.
EVI or NDVI, not both. The code does not handle more than that and weird
things will happen.

```{r}

fname = file.choose()

output_dir = dirname(fname)

raster = rast(fname)

#  some don't have it 
if (any(is.na(terra::time(raster)))) {
  raster = tbreak:::assign_time_to_raster(raster)
}

mm = minmax(raster)
if (any(mm[2,] < 1) && any(mm[1,] > -1)) {
  if (any(mm[1,] < -0.25)) {
    raster[raster < -0.25] = NA
  }
}


dim (raster)

time(raster)[1:6]

t1 = time(raster)[1]

plot (raster[[1]], main = as.character(t1))

#  initialise if needed
if (!exists("coord_string")) {
  coord_string = ""
}
if (!exists("bfast_h")) {
  bfast_h = 1/7
}
if (!exists("kml_filename")) {
  kml_filename = "one_pixel"
}
if (!exists("output_dir")) {
  output_dir = getwd()
}
if(!exists ("plot_title")) {
  plot_title = ""
}

get_plot_title = function (default) {
  response = svDialogs::dlg_input(
    message = "Enter a plot title", 
    default = default
  )
  plot_title = response$res
  if (length(plot_title) == 0) {
    plot_title = NULL
  }
  plot_title
}

```

#  Single cell BEAST result and plot

Generate a beast result for a single cell and plot it.

This is useful when the raster is too large to run the beast analysis
on.

Rerun this and the next few steps with different coords to explore your
data.

The coordinates to enter are obtained by right clicking on the map in
ArcGIS and choosing "Copy Coordinates".

In ArcGIS:

Make sure the map's display units uses either the MODIS Sinusoidal
coordinate system or WGS 1984. These can be checked and set under the
map's properties (right click on the map name in the table of contents).

The display units can also be changed by clicking the drop down to the
right of the coordinate at the bottom of the map pane.

```{r}

response = svDialogs::dlg_input(
  message = "Enter a coordinate from ArcGIS (MODIS or WGS84 coord sys)", 
  default = coord_string
)

resp = response$res
if (length (resp)) {
  coord_string = resp
  plot_title = get_plot_title(plot_title)
  
  ts_res = tbreak:::calc_and_plot_beast_modis_coord (raster, coord = coord_string, main = plot_title)
}


```


#  Single cell BFAST result and plot

Generate a bfast output for the same coord, or a different one if
preferred.

The "h" parameter is the segment size as a fraction of the time series.
Whole numbers are treated as days and converted to fractions internally.

```{r}

response = svDialogs::dlg_input(
  message = "Enter a coordinate from ArcGIS (MODIS or WGS84 coord sys)", 
  default = coord_string
)

resp = response$res
if (length (resp)) {
  coord_string = resp

  response = svDialogs::dlg_input(
    message = "BFAST h parameter (default is 1/7, whole numbers are treated as days)", 
    default = bfast_h
  )
  resp = response$res
  
  if (length(resp) > 0) {
    #  if we are a fraction of whole numbers
    if (all(grepl(pattern = r"(^\s*\d+/\d+\s*$)", x = resp))) {
      bfast_h = eval (parse (text = resp))
    } else {
      bfast_h = as.numeric(resp)
    }
    
    t = time(raster)
    if (bfast_h < 1) {
      message (
        sprintf(
          "H parameter of %s corresponds to %s days",
          bfast_h, 
          floor((t[length(t)] - t[1]) * bfast_h)
        )
      )
    }
  } else {
    h = 1/7
  }
  
  #message (bfast_h)
  plot_title = get_plot_title(plot_title)

  bfast_res = tbreak:::plot_bfast_modis_coord (raster, coord = coord_string, h=bfast_h, main = plot_title)
}


```

#  Single cell plain time series plot

Generate a plain time series plot for the same coord, or a different one
if preferred.

This is mostly useful for diagnostic checks as the bfast plot includes
the time series as the top plot.

```{r}

response = svDialogs::dlg_input(
  message = "Enter a coordinate from ArcGIS (MODIS or WGS84 coord sys)", 
  default = coord_string
)

resp = response$res
if (length (resp)) {
  coord_string = resp
  plot_title = get_plot_title(plot_title)
  ts_res = tbreak:::plot_ts_modis_coord (raster, coord = coord_string, main = plot_title)
}


```

# Create a KML polygon file of the cell associated with the coordinate

Export a polygon in KML format for the pixel corresponding with the specified coordinate.  This can then be loaded into a GIS or Google Earth.

```{r}

out_dir = svDialogs::dlgDir(title = "Choose the output directory", default = output_dir)$res
if (length (out_dir) > 0) {
  setwd(out_dir)
  output_dir = out_dir  #  save for re-use

  response = svDialogs::dlg_input(
    message = sprintf ("Enter a KML file name to save in %s", getwd()),
    default = kml_filename
  )
  
  col2hex = function(x, alpha = FALSE) {
    args = as.data.frame(t(col2rgb(x, alpha = alpha)))
    args = c(args, list(names = x, maxColorValue = 255))
    do.call(rgb, args)
  }
  
  resp = response$res
  if (length (resp)) {
    kml_filename = resp
    response = svDialogs::dlg_input(
      message = "Coordinate location",
      default = coord_string
    )
    coord_string = response$res
    if (length(coord_string)) {
      #  a limited set of colours at the moment (657 is too many...)
      col_choices = c("red", "orange", "yellow", "green", "blue", "violet", "white", "black")
      colour = col_choices[1]
      response = svDialogs::dlg_list(col_choices, title="Polygon outline colour")$res
      if (length(response)) {
        colour = response
      }

      poly = tbreak:::point_to_cell_polygon(coord_string, raster)
      poly = st_sf(poly)
      poly$OGR_STYLE = sprintf ("PEN(c:%s)", col2hex(colour))
      kml_filename = stri_replace_last(kml_filename, '', regex='\\.kml')
      write_sf(
        poly, 
        file.path(out_dir, sprintf ("%s.kml", kml_filename)),
        overwrite=TRUE
      )
    }
  }
}

```

#  Run BEAST on the whole raster 

Now run beast on the loaded raster. This might take a while.

Plotted results are the numbers of trend and season change points across the raster, plus the R^2 goodness of fit statistic for the model run across each cell.

```{r}


#  Try to gracefully handle errors without overwriting a
#  pre-existing beast_raster object if we fail.
#  That said, stopping the session kills it all anyway.
beast_result_tmp = tryCatch(
  {
    tbreak:::tiled_beast_modis (raster)
  }, error = function(e) {
    stop (sprintf ("Unable to process data\n%s", e))
  }
)

beast_result = beast_result_tmp
rm (beast_result_tmp)

beast_r2_raster = tbreak:::beastbit2raster(beast_result, "R2")
plot (beast_r2_raster, main = "R^2")

beast_ncps_raster = tbreak:::beastbit2raster(beast_result, "season", "ncp")
plot (beast_ncps_raster, main = "Number of season change points")

beast_ncpt_raster = tbreak:::beastbit2raster(beast_result, "trend", "ncp")
plot (beast_ncpt_raster, main = "Number of trend change points")



```

# Plot the BEAST result for one raster cell

Plot the result for one cell from a "beasted" raster.

This is most useful for when you want the exact same result that was generated in the preceding beast raster run.  Otherwise the single cell results are not entirely consistent.

```{r}


response = svDialogs::dlg_input(
  message = "Enter a coordinate from ArcGIS (MODIS or WGS84 coord sys)",
  default = coord_string
)

resp = response$res
if (length (resp)) {
  coord_string = resp
  plot_title = get_plot_title(plot_title)
  tbreak:::plot_beast_modis_coord (beast_result, coord = coord_string, main = plot_title)
}


```

#  Export the BEAST raster results

Now export the beast results to a set of rasters.

This generates many, many files.  You do not need all of them but see the help for the naming scheme.  

Look for the Value list under [the help system](https://cran.r-project.org/web/packages/Rbeast/refman/Rbeast.html#beast123).  Search for text "The output is a list object of class "beast" containing the following fields".


```{r}

out_dir = svDialogs::dlgDir(title = "Choose the output directory", default = output_dir)$res
if (length (out_dir) > 0) {
  setwd(out_dir)
  output_dir = out_dir

  output_prefix = svDialogs::dlg_input(message = "Filename prefix", default="")$res
  
  save_flag      = svDialogs::dlg_list  (choices = c("No", "Yes"), title = "Overwrite existing files?")$res
  overwrite_flag = ifelse (length(save_flag) > 0 && save_flag == "Yes", TRUE, FALSE)
  
  b_rasters = tbreak:::export_beast_rasters(beast_result, output_dir, output_prefix, overwrite_flag)
}

```

#  Run BFAST on the whole raster 

Now run BFAST on the loaded raster. This is slower than BEAST so will take a while.

Plotted results are the numbers of trend and season change points across the raster, plus the goodness of fit statistic (R^2).

BFAST is deterministic so to plot the results for a single cell you can use the block  above.  


```{r}

#  Try to gracefully handle errors without overwriting a
#  pre-existing beast_raster object if we fail.
#  That said, stopping the session kills it all anyway.
bfast_result_tmp = tryCatch(
  {
    tbreak:::bfast_raster (raster)
  }, error = function(e) {
    stop (sprintf ("Unable to process data\n%s", e))
  }
)

bfast_result = bfast_result_tmp
rm (bfast_result_tmp)

bfast_r2_raster = tbreak:::bfast_bit_to_raster(bfast_result, "R2")
plot (bfast_r2_raster, main = "R^2")

bfast_ncps_raster = tbreak:::bfast_bit_to_raster(bfast_result, "season_ncp")
plot (bfast_ncps_raster, main = "Number of season change points")

bfast_ncpt_raster = tbreak:::bfast_bit_to_raster(bfast_result, "trend_ncp")
plot (bfast_ncpt_raster, main = "Number of trend change points")


```


#  Export the BEAST raster results

Now export the beast results to a set of rasters.

This generates many, many files.  You do not need all of them but see the help for the naming scheme.  

Look for the Value list under [the help system](https://cran.r-project.org/web/packages/Rbeast/refman/Rbeast.html#beast123).  Search for text "The output is a list object of class "beast" containing the following fields".


```{r}

out_dir = svDialogs::dlgDir(title = "Choose the output directory", default = output_dir)$res
if (length (out_dir) > 0) {
  setwd(out_dir)
  output_dir = out_dir

  output_prefix = svDialogs::dlg_input(message = "Filename prefix", default="")$res
  
  save_flag      = svDialogs::dlg_list  (choices = c("No", "Yes"), title = "Overwrite existing files?")$res
  overwrite_flag = ifelse (length(save_flag) > 0 && save_flag == "Yes", TRUE, FALSE)
  
  b_rasters = tbreak:::export_bfast_rasters(bfast_result, output_dir, output_prefix, overwrite_flag)
}

```
