---
title: "BEAST with spatial data"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

Load needed libraries. Install if necessary.

```{r}

#  convoluted leftovers from initial version
#  should just load tbreak directly
pack_list = c("tbreak")
install_list = pack_list[!(pack_list %in% installed.packages()[,"Package"])]
if(length(install_list)) {
  library ("remotes")
  install_github("biogeospatial/tbreak")
}

for (p in pack_list) {
  library(p, character.only = TRUE)
}

#  Other libs needed below
pack_list = c("svDialogs", "terra")
install_list = pack_list[!(pack_list %in% installed.packages()[,"Package"])]
if(length(install_list)) {
  install.packages(install_list)
}
for (p in pack_list) {
  library(p, character.only = TRUE)
}


```

Now load the file we want and display some of its properties, and the
first band.

Make sure you load a geotiff or netcdf file with *ONLY ONE INDEX*, e.g.
EVI or NDVI, not both. The code does not handle more than that and weird
things will happen.

```{r}

fname = file.choose()

raster = rast(fname)

dim (raster)

time(raster)[1:6]

t1 = time(raster)[1]

plot (raster[[1]], main = as.character(t1))

```

Now run beast on the loaded data set. This might take a while.

```{r}

beast_result = tbreak:::beast_modis (raster)

plot (beast_result)



```

Some interaction. Rerun this step each time you want to plot a different
location.

The coordinates to enter are obtained by right clicking on the map in
ArcGIS and choosing "Copy Coordinates".

In ArcGIS:

Make sure the map's display uses the MODIS Sinusoidal coordinate system.
Also that the map's "Display units" are in "meters". Both of these can
be set under the map's properties (right click on the map name in the
table of contents). Display units is under the General section, the
Coordinate Systems should be self-documenting.

The display units can also be changed by clicking the drop down to the
right of the coordinate at the bottom of the map pane.

```{r}


response = svDialogs::dlg_input(
  message = "Enter a coordinate from ArcGIS (MODIS or WGS84 coord sys)"
)

coord_string = response$res

if (length (coord_string)) {
  tbreak:::plot_beast_modis_coord (beast_result, coord = coord_string)
}


```
Generate a bfast output for the same coord.

```{r}

response = svDialogs::dlg_input(
  message = "Enter a coordinate from ArcGIS (MODIS or WGS84 coord sys)", 
  default = coord_string
)

coord_string = response$res
if (length (coord_string)) {
#  tbreak:::plot_beast_modis_coord (beast_result, coord = coord_string)
}


```
Now export the beast results to a set of rasters.

```{r}

output_dir = svDialogs::dlgDir()$res

output_prefix = svDialogs::dlg_input(message = "Filename prefix", default="")$res

save_flag      = svDialogs::dlg_list  (choices = c("No", "Yes"), title = "Overwrite existing files?")$res
overwrite_flag = ifelse (length(save_flag) > 0 && save_flag == "Yes", TRUE, FALSE)

b_rasters = tbreak:::export_beast_rasters(beast_result, output_dir, output_prefix, overwrite_flag)


```

